---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "70%"
)
```
# netplot

The goal of netplot is to provide a more flexible yet nicer graph plotting function with better defaults

Some features:

1.  Auto-scaling of vertices using sizes relative to the plotting device.
2.  Embedded edge color mixer.
3.  True curved edges drawing.
4.  User-defined edge curvature.
5.  Nicer vertex frame color.
6.  Better use of space filling the plotting device.

## Installation

<!-- You can install the released version of netplot from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("netplot") -->
<!-- ``` -->

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("USCCANA/netplot")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(igraph)
library(netplot)
set.seed(1)
x <- sample_smallworld(1, 200, 5, 0.03)
l <- layout_with_fr(x)

plot(x, layout = l) # ala igraph
nplot(x, layout = l) # ala netplot
sna::gplot(intergraph::asNetwork(x), coord=l)
```


```{r}
nplot(
  x,
  edge.curvature = pi,
  edge.color.alpha = .2,
  edge.width.range = c(2,2),
  bg.col = "white",
  layout = l
  )
```

