---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "70%"
)
```
# netplot

[![Travis build status](https://travis-ci.org/USCCANA/netplot.svg?branch=master)](https://travis-ci.org/USCCANA/netplot)
[![Build status](https://ci.appveyor.com/api/projects/status/3k2m3oq6o99qcs0r?svg=true)](https://ci.appveyor.com/project/gvegayon/netplot)



An alternative graph visualization engine that puts an emphasis on 
  aesthetics at the same time of providing default parameters that provide visualizations
  that are out-of-the-box nice.

Some features:

1.  Auto-scaling of vertices using sizes relative to the plotting device.
2.  Embedded edge color mixer.
3.  True curved edges drawing.
4.  User-defined edge curvature.
5.  Nicer vertex frame color.
6.  Better use of space filling the plotting device.

## Installation

<!-- You can install the released version of netplot from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("netplot") -->
<!-- ``` -->

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("USCCANA/netplot")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(igraph)
library(netplot)
set.seed(1)
data("UKfaculty", package = "igraphdata")
l <- layout_with_fr(UKfaculty)

plot(UKfaculty, layout = l) # ala igraph
nplot(UKfaculty, layout = l) # ala netplot
sna::gplot(intergraph::asNetwork(UKfaculty), coord=l)
```


```{r}
nplot(
  UKfaculty,
  bg.col = "white",
  vertex.size.range = c(.01, .075, 5), 
  layout = l,
  edge.curvature = .5,
  edge.color.alpha = c(0, .5), edge.arrow.size = 0
  )
```

