---
title: "Graph drawing with netplot"
author: "George G. Vega Yon"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Graph drawing with netplot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Some features:

1.  Auto-scaling of vertices using sizes relative to the plotting device.
2.  Embedded edge color mixer.
3.  True curved edges drawing.
4.  User-defined edge curvature.
5.  Nicer vertex frame color.
6.  Better use of space filling the plotting device.


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

```{r pkgs-and-data, message=FALSE, warning=FALSE}
library(netplot)
library(igraph)
library(sna)
library(ggraph)

# We will use the UKfaculty data from the igraphdata package
data("UKfaculty", package = "igraphdata")

# With a fixed layout
set.seed(225)
l_ukf <- layout_with_fr(UKfaculty)
```

Since igraph and statnet do the plotting using base system graphics, we can't put everything in the same device right away. Fortunately, the `gridGraphics` R package allows us to reproduce base graphics using the `grid` system, which, in combination with the function `gridExtra::grid.arrange` will let us put both base and grid graphics in the same page. First me write a func

```{r}
# Function to map base to grid
map_base_to_grid <- function(fun) {
  gridGraphics::grid.echo(fun)
  grid::grid.grab()
}
```


```{r igraph-sna, fig.keep='none'}
# Mapping base to grid
ig <- map_base_to_grid(function() plot(UKfaculty, main = "igraph"))
nw <- map_base_to_grid(function() gplot(as.matrix(as_adj(UKfaculty)), coord = l_ukf, main = "sna"))
```

```{r nplot, fig.keep='none'}
nplot(UKfaculty, layout = l_ukf, main="netplot")
grid::grid.text("netplot", x = .05, y=.9)
np <- grid::grid.grab()
```

Here is an example with [`ggraph`](https://cran.r-project.org/package=ggraph)

```{r, fig.width=7, fig.height=5, fig.keep='none'}
ggraph::ggraph(UKfaculty, layout = l_ukf) +
  ggraph::geom_edge_arc() +
  ggraph::geom_node_point() +
  ggplot2::ggtitle("ggraph")

gg <- grid::grid.grab()
```


```{r, fig.width=7, fig.height=7, fig.cap="Comparision of default `igraph`, `sna`, `ggraph`, and `netplot` default call. `nplot` fills completely the plotting area, and adjusts vertex size, edge width, and edge arrows' size accordingly to the plotting area and plotting device."}
# Putting all together
gridExtra::grid.arrange(ig, nw, gg, np, nrow=2, ncol=2)
```


# Multiple plots

To arrange multiple plots in the same page we can use the `grid.arrange` function from the `gridExtra` package

```{r}
x_igraph  <- sample_smallworld(1, 20, 4, .05)
x_network <- intergraph::asNetwork(x_igraph)

l <- layout_nicely(x_igraph)

# Putting two plots in the same page (one using igraph and the other network)
gridExtra::grid.arrange(
  nplot(x_igraph, layout = l),
  nplot(x_network, layout = l), ncol=2, nrow=1
)
```


# Nodes

```{r node-scaling, fig.width=7, fig.height=7, fig.cap="Modifying `vertex.size.range`: Each figure shows a different parameter for the vertex size range. From left to right, and top down: (a) Has all vertices with the same scale of 2.5%, (b) "}

gridExtra::grid.arrange(
  nplot(UKfaculty, layout = l_ukf, vertex.size.range = c(.025, .025)),
  nplot(UKfaculty, layout = l_ukf, vertex.size.range = c(.01, .025)),
  nplot(UKfaculty, layout = l_ukf, vertex.size.range = c(.01, .025, 4)),
  nplot(UKfaculty, layout = l_ukf, vertex.size.range = c(.02, .05, 4)),
  ncol=2, nrow=2
)


```

```{r node-shapes, fig.width=4, fig.height=4, fig.cap="Number of sides for node drawing."}
nplot(make_ring(5), vertex.nsides = 2:6,
      vertex.size.range = c(.1,.1),
      edge.curvature = 0)
```

# Edges

```{r sim-net}
set.seed(12233)

x <- sample_smallworld(1, size = 10, nei = 2, .1)
l <- layout_with_fr(x)
```


```{r edge-color, fig.width=7, fig.height=3, fig.cap="Modifying `edge.color.mix`: Each figure shows a different parameter for the edge color mixer. From left to right, (a) colors the edges as alter, (b) mixes ego and alter's colors, and (c) only uses ego"}

gridExtra::grid.arrange(
  nplot(
    x, layout          = l,
    edge.color        = ~ego(mix=0, alpha = .1, col="black") + alter(mix=1),
    vertex.size.range = c(.05,.1)
    ),
  nplot(
    x, layout          = l,
    edge.color        = ~ego(mix=.5, alpha = .1, col="black") + alter(mix=.5),
    vertex.size.range = c(.05,.1)
    ),
  nplot(
    x, layout          = l,
    edge.color        = ~ego(mix=1, alpha = .1, col="black") + alter(mix=0),
    vertex.size.range = c(.05,.1)
    ),
  ncol=3, nrow=1
)

```

```{r edge-curvature, fig.width=7, fig.height=3, fig.cap="Modifying `edge.curvature`: Each figure shows a different parameter for the edge curvature. From left to right, (a) straight edges, (b) the edge between ego and alter is an arc that measures $\\pi/2$ radians (90 degree), and (c) the edge as an arc between ego and alter that measures $\\pi$ radians (180 degrees)."}
gridExtra::grid.arrange(
  nplot(x, layout = l, edge.curvature = 0, vertex.size.range = c(.05,.1)),
  nplot(x, layout = l, edge.curvature = pi/2, vertex.size.range = c(.05,.1)),
  nplot(x, layout = l, edge.curvature = pi, vertex.size.range = c(.05,.1)),
  ncol = 3, nrow=1
)
```

```{r edge-line-type, fig.width=6, fig.height=6, fig.cap="Changing the number of breaks in the edge (arc) and the type of line to be drawn."}
gridExtra::grid.arrange(
  nplot(x, layout = l, edge.line.breaks=3),
  nplot(x, layout = l, edge.line.lty = 2, edge.line.breaks=10),
  nplot(x, layout = l, edge.line.lty = 3, edge.line.breaks=10),
  nplot(x, layout = l, edge.line.lty = 4, edge.line.breaks=10),
  nrow=2, ncol=2
)
```

# Node labels

```{r}
data(UKfaculty, package = "igraphdata")
# fakenames <- sample
```

